### 2018-08-30 LatexNotes/Notes.tex
Added a bunch of files to the repo including the reading list (Literature), a scan of Andys original notes, my typed up Latex notes, my Julia codes for sampling from Gibbs distributions

\documentclass[12pt]{article}

% PACKAGES
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
\usepackage[margin=1in]{geometry}
%\usepackage[justification=RaggedRight]{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% LATEX DEFINITIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}

% Experimental parameters
\newcommand{\lamm}{\lambda^{-}}
\newcommand{\lamp}{\lambda^{+}}
\newcommand{\hm}{h^{-}}
\newcommand{\hp}{h^{+}}

% Theory for truncated KdV and stat mech
\newcommand{\uhat}{\hat{u}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Real}{\text{Re}}
\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
\newcommand{\Proj}{P_{\Lambda}}
\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
\newcommand{\sumn}{\sum_{n=1}^{N}}
\newcommand{\uDir}{u_{\text{Dir}}}
\newcommand{\Gibbs}{\mathcal{G}}

% Dashed integral
\def\Xint#1{\mathchoice
   {\XXint\displaystyle\textstyle{#1}}%
   {\XXint\textstyle\scriptstyle{#1}}%
   {\XXint\scriptstyle\scriptscriptstyle{#1}}%
   {\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
   \!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$}
     \vcenter{\hbox{$#2#3$}}\kern-.5\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}
\newcommand{\intt}{\dashint}%_0^{2 \pi}}
\newcommand{\dx}{\, dx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change}
\author{}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Mathematical Preliminaries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mathematical preliminaries}

\subsection{Truncated Fourier series}

We will be dealing with the truncated KdV equation (TKDV). 
Consider a real-valued function $u(x): [0, 2\pi] \to \RR$ given by a truncated Fourier series
\begin{equation}
u(x) = \sum_{k= -\Lambda}^{\Lambda} \uhat_k e^{i k x}	\qquad \text{for } 0 \le x \le 2\pi
\end{equation}
where
\begin{equation}
\uhat_k = \intt u(x) e^{-i k x} \dx
\end{equation}
Above, and throughout the paper, we use the notation
\begin{equation}
\intt \cdot \, \dx  = \frac{1}{2\pi} \int_0^{2 \pi} \cdot \, \dx
\end{equation}
We will be dealing with zero-momentum functions
\begin{equation}
M = \intt u(x) \dx = 0
\end{equation}
Let $\Fspace$ denote the space of all zero-momentum functions that can be represented by the above truncated Fourier Series (with maximum wavenumber $\Lambda$).

Since any element $u \in \Fspace$ is real valued and has zero momentum, it can be written as
\begin{equation}
u(x) = \sumk \uhat_k e^{i k x} + \uhat_k^{*} e^{-i k x} 
= 2 \Real \sumk \uhat_k e^{i k x}
\end{equation}
This implies that $\dim(\Fspace) = \Lambda$ as a complex vector space (or the dimension of the equivalent real vector space is $2\Lambda$).
For any $u \in \Fspace$, we introduce the energy
\begin{equation}
E = \frac{1}{2} \intt \Proj(u^2) \dx = \sumk \abs{\uhat_k}^2
\end{equation}
where $\Proj$ is the projection onto $\Fspace$.
In order to define the Hamiltonian for TKDV, we introduce the following components
\begin{equation}
H_3 = \frac{1}{6} \intt \Proj(u^3) \dx \qquad H_2 = \frac{1}{2} \intt \Proj(u_x^2) \dx
\end{equation}
We note that $H_3$ can be written explicitly in Fourier space as \cite{abramov2003hamiltonian} 
\begin{equation}
\label{H3dirsum}
H_3 = \frac{1}{6} \sum_{\substack{ k_1 + k_2 + k_3 = 0 \\ \abs{k_1}, \abs{k_2}, \abs{k_3} \le \Lambda}}\uhat_{k_1} \uhat_{k_2} \uhat_{k_3}
\end{equation}

\subsection{Some useful estimates/bounds for $H_2$ and $H_3$}

Since we examine the statistical mechanics of TKDV, it is valuable to have some a priori bounds and estimates for the components of the Hamiltonian, $H_2$ and $H_3$. In this section, we frequently appeal to the idea of an equipartition of energy among all the modes. For energy $E_0$, if the equipartition holds, then
\begin{equation}
\abs{u_k} = \sqrt{\frac{E_0}{\Lambda}} \qquad \text{for } k=1,\cdots,\Lambda
\end{equation}
We might expect equipartition of energy to hold, for example, in the case of long-time averages or ensemble averages (appealing to ergodicity).

\subsubsection{Expected value of $H_2$ given $E=E_0$}

First, we consider $H_2$. In Fourier space, $u_x$ is represented as
\begin{equation}
u_x(x) = 2 \Real \sumk i k \uhat_k e^{i k x}
\end{equation}
Thus, for equipartitioned energy, we have
\begin{equation}
H_2 = \sumk \abs{k \uhat_k}^2 = \frac{E_0}{\Lambda} \sumk k^2 = \frac{E_0}{6} (\Lambda+1)(2 \Lambda+1) \sim \frac{E_0 \Lambda^2}{3} \text{for } \Lambda \gg 1
\end{equation}
The above provides a useful estimate for the expected value of $H_2$ (to test the numerics for example).

\subsubsection{Upper bound on $H_3$ given $E=E_0$}

Next, consider $H_3$. By the symmetry of the cubic form, the expected value of $H_3$ is zero. However, to sample from Gibbs measures with non-vanishing inverse temperatures using an acceptance/rejection based algorithm, we will need to estimate the maximum allowed value of $H_3$ on the hypersphere of fixed energy $E = E_0$ so that the acceptance rate can be normalized. Note that the set $E=E_0$ is compact and therefore a maximum is guaranteed to exist. A reasonable guess for function that maximizes $H_3$ is something approximating a Dirac delta function, which naturally motivates the Dirichlet kernel
\begin{equation}
\label{Dkern}
\uDir(x) = \sqrt{\frac{E_0}{\Lambda}}\sumk e^{i k x}
\end{equation}
We can compute $H_3(\uDir)$ by using the direct summation formula \eqref{H3dirsum}. Since $\uhat_k = \sqrt{E_0/\Lambda}$, each term in this sum takes the value $(E_0/\Lambda)^{3/2}$, and so we simply need to count the number of terms present in this sum. A careful counting (being sure to dismiss any terms in which one of $k_1, k_2, k_3$ is zero or one of these terms exceeds $\Lambda$) gives
\begin{equation}
\label{H3Dir}
H_3(\uDir) = \frac{1}{2} E_0^{3/2} (\Lambda^{1/2} - \Lambda^{-1/2})
\end{equation}
Certainly, it must be that
\begin{equation}
\max H_3 \ge H_3(\uDir) 
\end{equation}
My conjecture is that
\begin{equation}
\max H_3 = H_3(\uDir) 
\end{equation}
or at least that this relationship holds asymptotically for $\Lambda \gg 1$.
Preliminary numerical tests seem to support this conjecture for tested values $\Lambda = 6, 8, 10, 20$. In fact, the maximum of $H_3$ encountered over about $10^6$ samples is usually significantly smaller than the value predicted in \eqref{H3Dir}. However, for $\Lambda = 4$, my tests encounter values of $H_3$ slightly larger than the predicted value.

Maximizing $H_3$ is perhaps more straightforward in physical space, where we discretize the interval $[0, 2 \pi)$ with 
\begin{equation}
x_n = \frac{2 \pi (n-1)}{N} \quad \text{for } n=1,\cdots,N
\end{equation}
where
\begin{equation}
N = 2 \Lambda \text{ or  } 2 \Lambda +1
\end{equation}
Both are valid choices for using standard real FFT packages.
%$x_n = 2 \pi (n-1)/N$ for $n=1,\cdots,N$, and $N$ is either $2 \Lambda$ or $2 \Lambda + 1$ (both are valid choices for using standard real FFT packages). 
In physical space, we have
\begin{equation}
\label{H3phys}
H_3 = \frac{1}{6} \intt \Proj (u^3) \dx \approx \frac{1}{6N} \sumn u_n^3
\end{equation}
The second relationship is only an approximation because de-aliasing would be required to make the right-hand side exactly equal to \eqref{H3dirsum}. Nonetheless, let us proceed with the estimate. For fixed energy, $E=E_0$, we require
\begin{equation}
E_0 = \intt \Proj(u^2) \dx = \frac{1}{N} \sumn u_n^2
\end{equation}
(I believe the above holds with no approximation being made due to the discrete Parseval identity)

Clearly, a maximizer of the right-hand-side of \eqref{H3phys} subject to $M=0$ and fixed $E=E_0$ is an approximate Dirac delta function --- i.e. a function with one extreme positive value, and the rest equal negative values,
\begin{align}
& w_1 = \sqrt{2 E_0 (N-1)}		\\
& w_n = -\sqrt{\frac{2 E_0}{N-1}}	\qquad \text{for } n \ne 1
\end{align}
This function is a maximizer because any slight modification (subject to keeping $M=0$ and $E=E_0$) would reduce the value of $H_3$.
Calculating the discrete Fourier transform of this function directly gives
\begin{align}
& \hat{w}_0 = 0	\\
& \hat{w}_k = \sqrt{\frac{2 E_0}{N-1}} \qquad \text{for } k \ne 1
\end{align}
Notice that if we choose $N = 2\Lambda + 1$, then our delta-function in physical space corresponds exactly to the Dirichlet kernel, $w(x) = \uDir(x)$. Now, if we calculate $H_3(w)$ purely in physical space using the approximation in \eqref{H3phys}, we get
\begin{equation}
H_3(w) = \frac{\sqrt{2} E_0^{3/2}}{3} \left( \frac{N-2}{\sqrt{N-1}} \right)
\end{equation}
Taking $N = 2\Lambda + 1$ gives
\begin{equation}
\label{H3physest}
H_3(w) = \frac{2 E_0^{3/2}}{3} \left( \Lambda^{1/2} - \Lambda^{-1/2} \right)
\end{equation}
So while \eqref{H3Dir} and \eqref{H3physest} take very similar forms, they differ by a multiplicative factor: 1/2 in Eq.~\eqref{H3Dir} versus 2/3 in Eq.~\eqref{H3physest}. Therefore, I suppose the error made by failing to de-alias in Eq.~\eqref{H3phys} is not negligible.

So the impasse here as follows: It is straightforward to maximize the approximate form of $H_3$ in physical space (the right-hand-side of \eqref{H3phys}), and justifying this maximization can be made completely rigorous in my mind. However, the approximation made by failing to di-alias is apparently not that small. On the other hand, we can calculate $H_3$ exactly in spectral space, but it is not clear how to maximize $H_3$ subject to the constraints --- I can only make a guess at the maximizer. To overcome this issue, one might naturally want to use the de-aliased form of $H_3$ in physical space (by taking $N = 4 \Lambda$ for example), but the problem here is that all discretized functions are not permissible. Only functions that correspond to a Fourier series truncated at wavenumber $\Lambda$ are permissible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TKDV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The truncated KdV equations}

We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC). The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$. Accordingly, the upstream and downstream Hamiltonians are given by
\begin{align}
\label{Hamiltonian}
& H^- = E_0^{1/2} H_3 - H_2 \\
& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Numerical Methods
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}

Here we describe the statistical numerical methods. The deterministic simulations of TKDV are described elsewhere. Below the microstate refers to the spectrum of displacement $\{ \uhat_k \}_{k=1}^{\Lambda}$.

\subsection{Sampling from microcanonical energy distribution}

We first consider sampling microstates from a microcanonical energy distribution with $E=E_0$ fixed and $M=0$ as always,
\begin{equation}
\Gibbs^0 = d\nu \propto \delta(E-E_0) \delta(M)
\end{equation}
This can be considered a canonical distribution in the Hamiltonian with zero inverse temperature (hence the notation $\Gibbs^0$). At times, it is easier to denote the distribution $d\nu$, so we will use both interchangeably. Note this is simply the uniform distribution on the hyper-sphere of $E=E_0$.

To sample from this distribution, we use the following algorithm
\begin{enumerate}
\item Sample each $\uhat_k$ for $k=1,\cdots,\Lambda$ from a normal distribution $N(0,1)$.
\item Compute $E = \sumk \abs{\uhat_k}^2$
\item Normalize each component by multiplying by $\sqrt{E_0/E}$.
\end{enumerate}
This algorithm was proven to converge to the measure $d\nu$ in \cite{abramov2003hamiltonian}. Throughout this paper, we normalize the energy for the sampling distribution to unity. This is the reason why $E_0$ appears in the Hamiltonian in \eqref{Hamiltonian}. Hence, we will always be sampling from
\begin{equation}
\Gibbs^0 \propto \delta(E-1) \delta(M)
\end{equation}

\subsection{Acceptance/rejection algorithm to sample from a mixed microcanonical/canonical distribution.}

Now consider the problem of sampling from a distribution that is microcanonical in energy and canonical in the Hamiltonian
\begin{equation}
\Gibbs^{\pm} \propto \delta(E-1) \delta(M) e^{-\theta^{\pm} H^{\pm}}
\end{equation}
Here, the Hamiltonian can be either the upstream or downstream, $H = H^-$ or $H^+$ and the inverse temperature $\theta$ can also take different values upstream and downstream, $\theta^-$ and $\theta^+$. To sample from this distribution, we use the following acceptance/rejection algorithm.
\begin{enumerate}
\item Sample a large number of microstates from $\Gibbs^0$.
\item For each, compute $H_3$ and $H_2$. Note: this is the most expensive step in the algorithm. We will therefore store all the values of $H_3$ and $H_2$ so they need not be recomputed.
\item For each sample, compute $H^+$ and $H^-$.
\item Assign a provisional acceptance rate as $e^{- \theta H}$ (for either upstream or downstream). The only issue is this acceptance rate is not yet normalized.
\item Find the maximum of the acceptance rate over all samples $m = \max{e^{-\theta H}}$.
\item Use this value to normalize the acceptance rates, $e^{- \theta H}/m$.
\item Decide to accept or reject each microstate by picking a uniform variable $v \in [0,1]$ and testing if $e^{- \theta H}/m < v$.
\item Store the microstates, as well as the values of $H_3$ and $H_2$, that have been accepted.
\end{enumerate}
The above is a conceptual description of the algorithm. In practice, it is implemented in a way to be computational efficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminary results}
We show some preliminary results below. 
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist1}
\caption{Left: Sampling $H^-$ from the incoming Gibbs distribution $G^-$ with three different inverse temperatures $\theta^-$ (see legend). Right: Sampling the $H^+$ from an outgoing distribution $G^+$ with the same inverse temperatures.
In these tests, $\Lambda = 10$, $E_0 = 1$, $\lambda = 1$, $D_0 = 0.6$, and the number of samples is $1 \times 10^7$.}
\label{Hist1}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist2}
\caption{Sensitivity to changes in $D_0$.
Left: Downstream histogram with $D_0 = 0.6$ (same as in Fig.~\ref{Hist1}). Right: Same with $D_0 = 0.4$. Notice the histogram changes significantly.
}
\label{Hist2}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
%Notice that even though the depth ratio changes only slightly, the histograms change significantly. The upstream histograms do not depend on $D_0$ and are therefore not shown.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist3}
\caption{Sensitivity to changes in the energy $E_0$.
Left: Downstream histogram with $E_0 = 1$ (same as in Fig.~\ref{Hist1}). Right: Same with $E_0 = 5$. Notice a slight change in the histogram.}
\label{Hist3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
 %Notice that the energy changes substantially, but the histograms change only slightly. While the upstream histogram ostensible depends on $E_0$, the changes are nearly imperceptible and therefore not shown.

  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microup0}
\caption{Sampling the upstream microstates with $\theta^- = 0$ (and $E_0 = 4$). Left: histogram of the surface displacement, $u^-$. Right: ensemble average of the spectrum, $\mean{\abs{\uhat_k}}$. Notice symmetric distribution of the displacement and the equipartition of energy amongst modes.}
\label{microup0}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microup3}
\caption{The upstream microstates with $\theta^- = -0.3$ (and $E_0 = 4$ as before). Notice the distinct tilt to the spectrum due to $\theta^- < 0$.}
\label{microup3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microdn3}
\caption{Examination of the downstream microstates with $\theta^+ = -0.3$ and $D_0 = 0.5$ ($E_0 = 4$ as before). Notice the skewness of $u$!!! The spectrum is tilted, though not as much as it is upstream.}
\label{microdn3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

\newpage
\newpage
\section{Experimental parameters}

Some relevant parameters from the experiment are given in the table.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Some relevant experimental parameters} 
\vspace{0.3 pc}
\begin{tabular}{c l l}
\hline
\hspace{0.5pc} Symbol
\hspace{0.5pc} & Meaning 
\hspace{0.5pc} & Value \\
\hline
N/A		& Length of the experimental tank (in wave direction)	& 6 m	\\
$L_0$	& Width of the experimental tank	 (transverse)		& $L_0 = $ 20 cm	\\
$f_0$	& Dominant forcing frequency in experiments	& $f_0 = $ 2 Hz	\\
$\hm$	& Depth on deep side						& $\hm = 12.5$ cm \\
$\hp$	& Depth on shallow side					& $\hp = 3$ cm \\
$\lamm$	& Wavelength induced by 2 Hz forcing on deep side	& $\lamm = 38$ cm	\\
$\lamp$	& Wavelength induced by 2 Hz forcing on shallow side	& $\lamp = 25$ cm	\\
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

\begin{comment}
% Majda's notes typed up
\section{Deterministic model for waves over abrupt depth change}

The surface-displacement is denoted $u(x,t)$.

Modeling assumptions
\begin{enumerate}
\item Waves in a narrow channel (width $L_0$) are nearly unidirectional and $2 \pi L_0$ periodic (purely a mathematical idealization, not realistic I believe). Where $L_0 < \lambda$, with $\lambda$ a characteristic wavelength. As seen in the table, this later assumption is reasonable since $L_0 / \lamm \approx 0.5$
\item The role of the paddle forcing is to generate a state with `zero momentum', $M = \intt u \, dx = 0$ (this assumption is strongly supported by experimental measurements pointwise!) and with some given energy $E_0 = \frac{1}{2} \intt u^2 \, dx$ (not an assumption, just a statement). Here, $E_0$ is a function of the strength of the paddle forcing.
\item Assume a Galerkin truncation of $u(x,t) \approx u_{\Lambda}(x,t)$, where $\Lambda$ (fixed) will probably be 20 or 30.
\item The `time' coordinate $t$  corresponds physically to a downstream spatial coordinate (i.e.~$X \approx \eps x$ in Johnson's notation). The abrupt depth change (ADC) corresponds to $t = t_0$ or $T_{ADC}$.
\end{enumerate}
\end{comment}

\bibliographystyle{plain}
\bibliography{Notesbib}

\end{document}